---
title: "Introduction to the ngramr.plus R package"
author: "David Brown"
output: rmarkdown::html_vignette
bibliography: ngramr_bib.bib
link-citations: yes
nocite: |
  @*
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

## Load the ngramr.plus package

Load the package, as well as others that we'll use in this vignette.

```{r setup, message = FALSE, error=FALSE}
library(ngramr.plus)
library(tidyverse)
```

## Retrieving data

Running the functions is straightforward, but remember that it can take a couple minutes if your are loading one of the larger Google Books data tables. Here we return counts of *zinger* by year from the data tables for US English.

```{r xray_year, message=FALSE, results='hide'}
z_year <- google_ngram(word_forms = "zinger", variety = "us", by = "year")
```

Check the data:

```{r echo=FALSE}
knitr::kable(head(z_year))
```

In current usage, *zinger* denotes a kind of cutting quip. Its [supposed origin](https://www.etymonline.com/word/zinger) is as a baseball term that becomes generalized in the middle of the 20th century. Does this explanation comport with the data?

## Plot the data

To partly answer such a question, we can plot the data:

```{r year_plot, fig.height=4.5, fig.width=7}
  ggplot(z_year %>% filter(Year > 1799), aes(x=Year, y=Per_10.6)) +
    geom_point(size = .5) +
    geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), size=.25) +
    labs(x="year", y = "frequency (per million words)")+ 
    theme(panel.grid.minor.x=element_blank(),
          panel.grid.major.x=element_blank()) +
    theme(panel.grid.minor.y =   element_blank(),
          panel.grid.major.y =   element_line(colour = "gray",size=0.25)) +
    theme(rect = element_blank()) +
    theme(legend.title=element_blank()) +
    theme(axis.title = element_text(family = "Arial", color="#666666", face="bold", size=10))
```

There does appear to be some circulation prior to the mid-20th century. To explain that, we can look at [the underlying texts in Google Books](https://www.google.com/search?q=%22zinger%22&tbm=bks&tbs=cdr:1,cd_min:1800,cd_max:1893&lr=lang_en).

## By decade

Next, we'll return counts of *zinger* and *zingers* by decade from the data tables for British English.

```{r xray_decade, message=FALSE, results='hide'}
z_decade <- google_ngram(word_forms = c("zinger", "zingers"), variety = "gb", by = "decade")
```

Check the data:

```{r echo=FALSE}
knitr::kable(head(z_decade))
```

## Ploting the data

Now we can filter and plot our by-decade data:

```{r decade_plot, fig.height=4.5, fig.width=7}
ggplot(z_decade %>% filter(Decade > 1799), aes(x=Decade, y=Per_10.6)) +
  geom_bar(stat = "identity") +
  labs(x="decade", y = "frequency (per million words)")+ 
  theme(panel.grid.minor.x=element_blank(),
         panel.grid.major.x=element_blank()) +
  theme(panel.grid.minor.y =   element_blank(),
        panel.grid.major.y =   element_line(colour = "gray",size=0.25)) +
  theme(rect = element_blank()) +
  theme(legend.title=element_blank()) +
  theme(axis.title = element_text(family = "Arial", color="#666666", face="bold", size=10))
```


## Bibliography


